---
title: "helminthR"
author: "Tao Huang"
date: "October 22, 2018"
output: html_document
---
#install and load the package
```{r}
#devtools::install_github("rOpenSci/helminthR")
library("helminthR")
library("ggmap")
library("googleway")
```

###load and clean the locations data
```{r}
locations
data(locations)
r_cty<-c("East Coast","Central","Northern Europe","Western Europe","Eastern & Central Europe","Southeast Europe","Southwest Europe","South","North West","South West","Northern Territory","North Island","South Island","Lesser Islands","Polar Sea","Greenland/Norwegian zone","East Greenland coast","Arctic Ocean","Marine - all records","Marine - no area specified","Icelandic coast","Northern Norwegian coast","USSR zone","Alaskan/Canadian zone","Bering Sea","Atlantic Ocean","NW Atlantic","Gulf of Mexico","NE South American coast","Coast of USA","Canadian coast","Brazilian coast","Argentine/Uruguayan coast","Falkland coast","Australian coast","Coast of N Territories","Coast of Queensland","Coast of NSW","Coast of Victoria","Coast of Tasmania","Coast of Indonesia & Malaysia","West Australian coast","South Austrlian coast","Coast of East Africa","South African coast","Coast of Madagascar","Atlantic Zone","Indian Ocean Zone","Pacific Ocean Zone"
         )
locations<-locations[!(locations$Location %in% r_cty),]
```

###get the table of Host-Parasite-Location
```{r}
for (cty in locations$Location)
{

f<- helminthR::findLocation(location=cty
                                     )
 f$Location<-cty

  assign(paste("HostPars_",cty,sep = ""),f )

}
l.df <- lapply(ls(pattern="HostPars_"), function(x) get(x))

l_df<-do.call("rbind", l.df)

write.csv(l_df,file="l_df.csv")

```

###the zooscore_trait list (n=748)
```{r}
for_z<-read.csv("C:/Users/huangt/Documents/Zooscores/Zooscore_trait_Helminth.csv",header=T,sep=',')
head(for_z)
l_df_z<-l_df[l_df$Parasite %in% for_z$ParasiteCorrectedName_Zooscores_VR_Ver5.0_Final,]
```

#read the Host-Parasite-Country table
```{r}
l<-read.csv("C:/Users/huangt/Documents/helminth_data/l.df.csv",header=T,sep=',')
head(l)
```

###to get the spp. in helminthR database
```{r}
dim(l)
#unique species
u_l<-unique(l$Parasite)
length(u_l)
z_in_l<-for_z[for_z$Match %in% u_l,]
dim(z_in_l)
z_not_l<-for_z[!(for_z$Match %in% u_l),]
dim(z_not_l)
```

###MERGE THE ZOOSCORE AND HELMINTHR
```{r}
z_in_l_w_l<-merge(z_in_l,l,by.x="Match",by.y="Parasite")
 dim(z_in_l)
#595  53
 dim(l)
#446532      5
dim(z_in_l_w_l)
#62562    57
```

###CREATE THE UNIQUE COMB OF PARASITE AND COUNTRY
```{r}
l$comb<-paste(l$Parasite,"_",l$Country,sep='')
head(l)
dim(l)
u_comb<-unique(l$comb)
length(u_comb)
library(dplyr)
library(tidyr)
u_comb<-as.data.frame(u_comb)
dim(u_comb)
write.csv(u_comb,file="u_comb.csv")
#comb %>% separate(comb, c("Parasite","Country"), sep="_")

```

```{r}
#separate the column in excel
#save as helminth_country.csv
```

#subset unique country-helminth pair to the helminth we have (n=13587)
#check_na<-as.data.frame(table(h_c_z$PARASITE)) no NAs
#check_na<-as.data.frame(table(h_c_z$COUNTRY)) no NAs
```{r}
h_c<-read.csv("helminth_country.csv",header=T,sep=',')
head(h_c)
dim(h_c)
#154652      3

length(unique(h_c$PARASITE))
for_z<-read.csv("C:/Users/huangt/Documents/Zooscores/Zooscore_trait_Helminth.csv",header=T,sep=',')
head(for_z)
h_c_z<-h_c[h_c$PARASITE%in%for_z$Match==T,]
dim(h_c_z)
length(unique(h_c_z$PARASITE))
write.csv(h_c_z,file="h_c_z.csv")
```

###INDEX OF PARASITE WITH COUNTRY NAMES IN THE SHP (n=4510) (idx=a)
```{r}
#country.dbf
#save it as country.csv
cnt<-read.csv("cntry00/country.csv",header=T,sep=',')
head(cnt)
#locations are country names
a<-which(h_c_z$COUNTRY%in%cnt$CNTRY_NAME==T)
h_c_z$cate<-" "
h_c_z[a,]$cate<-"a"
head(h_c_z)
```

###INDEX OF PARASITE WITH location NAMES not IN THE SHP (n=9077) (idx=b)
#check_na<-as.data.frame(table(h_c_z_n$COUNTRY)) no NAs
```{r}
b<-which(h_c_z$COUNTRY%in%cnt$CNTRY_NAME==F)
h_c_z[b,]$cate<-" "
h_c_z_n<-h_c_z[b,]
dim(h_c_z_n)
```

###USE geocode to get the lat/lon of these countries
```{r}
#list of locations that are not country names
n_list<-unique(h_c_z_n$COUNTRY)
n_list_df<-as.data.frame(n_list)

n_list_g<-geocode(as.character(n_list), source = "dsk")

n_list_df$lat<-" "
n_list_df$lon<-" "

n_list_df$lat<-n_list_g$lat
n_list_df$lon<-n_list_g$lon

length(n_list)
dim(n_list_g)
dim(n_list_df)

head(n_list)
head(n_list_g)
head(n_list_df)
#write.csv(n_list_df,file="n_list_df.csv")
```

#index the locations which can be extracted from geocode (idx=c)
```{r}
n_list_df_rm<-n_list_df[is.na(n_list_df$lat)==F,]
head(n_list_df)
head(n_list_df_rm)
dim(n_list_df)
dim(n_list_df_rm)
c<-which(h_c_z$COUNTRY%in%n_list_df_rm$n_list==T)
length(c)
h_c_z[c,]$cate<-"c"
#write.csv(n_list_df_rm,file="n_list_df_rm.csv")
```

#index the locations which can't be extracted from geocode (idx=dd)
```{r}
d<-which(b%in%c==F)
dd<-b[d]
h_c_z[dd,]$cate<-"dd"
#b: 58   59 60 61   62   63   64   65   67   68   69
#c: 58   59 60 67   68   69
#dd: 61   62   63   64   65 
```

#index c overlap with country shp  
```{r}
dim(h_c_z[c,])
#write.csv(h_c_z[c,],file="h_c_z_idxc.csv")
#to get the lat/lon for h_c_z
h_c_z_m<-merge(h_c_z,n_list_df,by.x="COUNTRY",by.y="n_list")
dim(h_c_z_m[is.na(h_c_z_m$lat)==T,]) #index dd
dim(h_c_z)
dim(h_c_z_m) #9077
head(h_c_z_m)
head(n_list_df)

write.csv(h_c_z_m[h_c_z_m$cate=="c",],file="h_c_z_m_idxc.csv") 


```

#idx e &ce
```{r}
#h_c_z_m_idxc_c.shp
library(foreign)
h_c_z_m_idxc_c_t<-read.dbf("h_c_z_m_idxc_c_t.dbf")
h_c_z[h_c_z$COUNTRY%in%h_c_z_m_idxc_c_t$COUNTRY==T,]$cate<-"e"
h_c_z[h_c_z$cate=="c",]$cate<-"ce"
```

```{r}
#spatial join to get the country names (h_c_z_m_idxc_c_j.shp)
h_c_z_m_idxc_c_j_t<-read.dbf("h_c_z_m_idxc_c_j_t.dbf")
h_c_z_m_idxc_c_j_t_u<-unique(h_c_z_m_idxc_c_j_t[c("COUNTRY", "CNTRY_NAME")])
h_c_z2<-merge(h_c_z,h_c_z_m_idxc_c_j_t_u,by.x="COUNTRY",by.y="COUNTRY",all.x = TRUE)
head(h_c_z2)
dim(h_c_z2)
h_c_z2[h_c_z2$cate=="e",]$country_shp<-as.character(h_c_z2[h_c_z2$cate=="e",]$CNTRY_NAME)
head(h_c_z2,21)
```

#Ania's list
```{r}
gg_cant<-read.csv("no_loc_AM_lookup.csv",header=T,sep=',')
head(gg_cant)
```

#idx g
```{r}
#C:\Users\huangt\Documents\no_loc_AM_lookup_c_t.dbf intersect w country
no_loc_AM_lookup_c_t<-read.dbf("no_loc_AM_lookup_c_t.dbf")
h_c_z2[h_c_z2$COUNTRY%in%no_loc_AM_lookup_c_t$Location==T,]$cate<-"f"
h_c_z2[h_c_z2$cate=="dd",]$cate<-"g"
```
#country shp names for index f (idx=f)
```{r}
#spatial join to get the country names (no_loc_AM_lookup_c_j.shp)
no_loc_AM_lookup_c_j_t<-read.dbf("no_loc_AM_lookup_c_j_t.dbf")
no_loc_AM_lookup_c_j_t_u<-unique(no_loc_AM_lookup_c_j_t[c("Location", "CNTRY_NAME")])
h_c_z3<-merge(h_c_z2,no_loc_AM_lookup_c_j_t_u,by.x="COUNTRY",by.y="Location",all.x = TRUE)
head(h_c_z3)
dim(h_c_z3)
h_c_z3[h_c_z3$cate=="f",]$country_shp<-as.character(h_c_z3[h_c_z3$cate=="f",]$CNTRY_NAME.y)
head(h_c_z3)
```
#COUNT NUMBER OF HELMINTHS IN EACH COUNTRY
```{r}
#unique country shp and parasite comb
h_c_z3_u<-unique(h_c_z3[c("country_shp", "PARASITE")])
dim(h_c_z3)
dim(h_c_z3_u)
count<-as.data.frame(table(h_c_z3_u$country_shp))
write.csv(count,file="count_by_country.csv")
```

```{r}
h_c_z$country_shp<-""
h_c_z[h_c_z$cate=="a",]$country_shp<-as.character(h_c_z[h_c_z$cate=="a",]$COUNTRY)

```


#the R package can find the lat and long but it doesn't belong to any country
```{r}
head(n_list_df_rm_c_t[n_list_df_rm_c_t$CNTRY_NAME=="",])
#the shp for this is n_list_df_rm_c2.shp
```

#remove water bodies
```{r}
library(data.table)
nnnn<-n_list_df_rm_c_t[n_list_df_rm_c_t$CNTRY_NAME=="",]
nnnnnn<-nnnn[!(nnnn$n_list %like% "Sea"),]
no_Antilles<-nnnnnn[!(nnnnnn$n_list%like%"Antilles"),]
no_ocean<-no_Antilles[!(no_Antilles$n_list%like%"Ocean"),]
no_bay<-no_ocean[!(no_ocean$n_list%like%"Bay"),]
no_island<-no_bay[!(no_bay$n_list%like%"Island"),]
no_med<-no_island[!(no_island$n_list%like%"Mediterranean"),]
no_channel<-no_med[!(no_med$n_list%like%"English Channel"),]
no_Gulf<-no_channel[!(no_channel$n_list%like%"Gulf"),]
dim(no_channel)
head(no_channel)
tail(no_Gulf)
```

#h_c_z IS THE HELMINTH(WITH ZOOSCORE)-COUNTRY TABLE
```{r}
h_c_z_c2<-merge(h_c_z,n_list_df_rm_c_t,by.x="COUNTRY",by.y="n_list",all=T)
head(n_list_df_rm_c_t)
dim(n_list_df_rm_c_t)
head(h_c_z)
dim(h_c_z)
head(h_c_z_c2)
dim(h_c_z_c2)
```

```{r}
h_c_z_c3<-h_c_z_c2
h_c_z_c3[is.na(h_c_z_c2$CNTRY_NAME)==F,]$COUNTRY<-h_c_z_c3[is.na(h_c_z_c2$CNTRY_NAME)==F,]$CNTRY_NAME
head(h_c_z_c3)
dim(h_c_z_c3)
#is.na(h_c_z_c2$CNTRY_NAME)==F
```

#find the unique county-helminth pair after updating locations
```{r}
head(h_c_z_c3)
h_c_z_c3$pair<-paste(h_c_z_c3$COUNTRY,"_",h_c_z_c3$PARASITE,sep='')
head(h_c_z_c3[!(h_c_z_c3$COUNTRY==""),])
write.csv(unique(h_c_z_c3[!(h_c_z_c3$COUNTRY==""),]$pair),file="new_pairs.csv")
```

```{r}


#save as new_pair2.csv

```

#COUNT HELMINTH NUMBERS FOR EACH COUNTRY
```{r}
new_pairs2<-read.csv("new_pairs2.csv",header=T,sep=',')
new_pairs2$count<-1
new_agg<-aggregate(new_pairs2$count, by=list(Category=new_pairs2$COUNTRY), FUN=sum)
head(new_agg)
write.csv(new_agg,file="new_agg.csv")
```


```{r}

h_c_z_cc3<-aggregate(h_c_z_c3$count, by=list(Category=h_c_z_c3$COUNTRY), FUN=sum)

head(h_c_z_cc3)
write.csv(h_c_z_cc3,file="h_c_z_cc3.csv")
```


```{r}
#count<-as.data.frame(table(n_list_g$lon))
#count<-count[order(-count$Freq),]
dim(n_list_g[is.na(n_list_g$lon)==T,])
#39  2
no_loc<-n_list_g[is.na(n_list_g$lon)==T,]
head(no_loc)
```

```{r}
i<-which(is.na(n_list_g$lon)==T)
n_list_i<-n_list[i]
write.csv(n_list_i,file="no_loc.csv")
```


