---
title: "helminthR"
author: "Tao Huang"
output: html_document
---

#install and load the package
```{r}
#devtools::install_github("rOpenSci/helminthR")
#devtools::install_github("dkahle/ggmap")
#install.packages("ggmap")
#install.packages("googleway")
#install.packages("Rcpp")
library("Rcpp")
library("ggmap")
library("helminthR")
library("foreign")
library("googleway")
library(dplyr)

```




###the zooscore_trait list (n=748)
```{r}
l_df<-read.csv("C:/Users/huangt/Documents/Mapping_helminths-master/l.df.csv",header=T,sep=',')
for_z<-read.csv("C:/Users/huangt/Documents/Mapping_helminths-master/Zooscore_trait_Helminth_201904.csv",header=T,sep=',')
head(for_z)
l_df_z<-l_df[l_df$Parasite %in% for_z$ParasiteCorrectedName_Zooscores_VR_Ver5.0_Final,]
```

###to get the unique helminth spp. in helminthR database
```{r}
l<-l_df
dim(l)
#unique species
u_l<-unique(l$Parasite)
length(u_l)
z_in_l<-for_z[for_z$Match %in% u_l,]
dim(z_in_l)
z_not_l<-for_z[!(for_z$Match %in% u_l),]
dim(z_not_l)
```

###MERGE THE ZOOSCORE AND HELMINTHR
```{r}
z_in_l_w_l<-merge(z_in_l,l,by.x="Match",by.y="Parasite")
 dim(z_in_l)
#595  53
 dim(l)
#446532      5
dim(z_in_l_w_l)
#62562    57
```

###CREATE THE UNIQUE COMB OF PARASITE AND COUNTRY
```{r}
helminth_country<-unique(l[,c("Parasite","Country")])


```


#subset unique country-helminth pair to the helminth we have (n=13587)
```{r}

h_c<-helminth_country
head(h_c)
dim(h_c)

length(unique(h_c$Parasite))

h_c_z<-h_c[h_c$Parasite%in%for_z$Match==T,]
dim(h_c_z)
length(unique(h_c_z$Parasite ))

```

###USE geocode to get the lat/lon
```{r}
#list of locations that are not country names
h_c_z_loc<-unique(h_c_z$Country )
h_c_z_loc<-as.data.frame(h_c_z_loc)
register_google(key = "your key")


has_google_key()

h_c_z_loc_g<-geocode(as.character((h_c_z_loc$h_c_z_loc ))
                     #, source = "dsk"
                     )
h_c_z_loc$lat<-" "
h_c_z_loc$lon<-" "

h_c_z_loc$lat<-h_c_z_loc_g$lat
h_c_z_loc$lon<-h_c_z_loc_g$lon
```

```{r}
write.csv(h_c_z_loc,file="C:/Users/huangt/Documents/Mapping_helminths-master/h_c_z_loc.csv")
write.csv(h_c_z_loc[!is.na(h_c_z_loc$lat),] ,file="C:/Users/huangt/Documents/Mapping_helminths-master/h_c_z_loc_rm_NA.csv")
```
#geocode can't find them
```{r}
write.csv(h_c_z_loc[is.na(h_c_z_loc$lat),] ,file="C:/Users/huangt/Documents/Mapping_helminths-master/location_NA.csv")
```

#geocode can't find them 
```{r}
no_loc_AM<-read.csv("C:/Users/huangt/Documents/Mapping_helminths-master/no_loc_AM lookup.csv",header=T,sep=',')
missing_loc<-merge(no_loc_AM,h_c_z_loc[is.na(h_c_z_loc$lat),],by.x= "Location", by.y= "h_c_z_loc" , all=T)
write.csv(missing_loc[is.na(missing_loc$Latitude)==T,] ,file="C:/Users/huangt/Documents/Mapping_helminths-master/missing_loc.csv")

```


#add a col ecoregion
```{r}
H_C_Z_LOC_RM_na_eco<-read.dbf("C:/Users/huangt/Documents/Mapping_helminths-master/H_C_Z_LOC_RM_na_eco.dbf")
```

#add lat/lon to helminth-country
```{r}
h_c_l<-merge(h_c_z,H_C_Z_LOC_RM_na_eco,by.x= "Country", by.y="h_c_z_loc",all=T )
dim(h_c_z)
dim(h_c_z_loc)
dim(h_c_l)
```

#count ecoregions of each helminth
```{r}
h_c_l_u<-unique(h_c_l[,c( "Parasite",  "FID_2" )])
h_c_l$count<-1
h_c_l_agg<-aggregate(h_c_l$count, by=list(Category=h_c_l$Parasite), FUN=sum)
head(h_c_l_agg)

colnames(h_c_l_agg)<-c("Parasite",  "Number_of_ecoregion")
write.csv(h_c_l_agg,"C:/Users/huangt/Documents/Mapping_helminths-master/Helminth_number_of_ecoregion.csv")

```
